---
title: "Report"
author: "Team IG"
date: "11/15/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,include=FALSE}
pacman::p_load(readxl,dplyr,tidyverse,ggplot2,plotly,ggpubr,MASS,fitdistrplus,mediation,geepack,conf,car,knitr,gclus,scatterplot3d, ppcor)
```

```{r,include=FALSE}
data<-read_excel("Data for the Final Project.xlsx")
data<-as_tibble(data)
#select variables that needed for the project and rename some variables
data2 <- data %>% 
  dplyr::select(State,Highschool,`Union density`,GDP,Urbanity, Population, Export,`Unemployment rate`,`Cost of living`,Gini, White, Black, Hispanic)
names(data2)[3]<-"Union_density"
names(data2)[8]<-"Unemployment_rate"
names(data2)[9]<-"Cost_of_living"
#data transformation: generate a new variable called GV to measure the level of racial diversity
data3 <- data2 %>% 
  mutate(Racial_diversity=1-White^2-Black^2-Hispanic^2) %>%
  mutate(norm_Racial_diversity = Racial_diversity*(3/2))
```

# Introduction
## VOC theory 
\qquad VoC theory (Varieties of Capitalism) literature has been a classic theory through out these decades and it mainly includes two prominent models which are Liberal Market Economies (LMEs) and Coordinated Market Economies (CMEs). This theory has been developing significantly and helping securing a firm root in the field of political economy. LME which exemplified by US, UK, the characteristic of this model is  Meanwhile, the CME represented by Sweden and Denmark. The difference between the LME and CME is that the profits of a company mainly comes from the supply and demands in LME, while, the companies in the CME are basically shape the form a core competitiveness through non market relationships.  

\qquad On the country scale of United States, it would be be classified as Liberal Market Economies while, while in the state scale, it likely to be a CME model. 

## Research question
\qquad According to the VoC theory, the United States is defined as liberal market economies and one characteristic of this model is that the union density is proportional to the vocational training level. While, the previous model contains 9 variables from dataset and defined as a linear model shows below:  
```{r,echo = F}
fittest<-lm(Highschool~Union_density+GDP+Urbanity+Population+Export+Unemployment_rate+Cost_of_living+Gini,data = data3)
```
\qquad The summary of the model shows to the coefficient, p-value, and other useful information.  
```{r,echo = F}
summary(fittest)
```
\qquad In the result, the coefficient of the high school is negative which means if union density arises, the high school would correspondingly decrease. This phenomenon apparently conflicts with properties of LME in VoC theory. A potential explaination comes that racial diversity in each state might leads to this result.

## Literature review
\qquad Market-baised management: Market-Based Management (MBM) enables organizations to succeed in the long term by applying the principles that allow free societies to prosper. Just as upholding values such as free speech, property rights, and progress is important to a healthy, growing society, it is also pivotal in fostering a healthy, growing organization.  

\qquad Nation-level/state level: US is called LME because its national governing framework does not has institutions that aggregate preference for public goods, such as high-skill development.

## Dataset interpretation
**Descriptions: **
Dataset details:
-51 observations(51 states in US) and 31 variables.
-Dependent variables included in the model: high school (represent vocational level).
-Independent variables included in the model: union density, racial(white/black/hispanic), GDP, Urbanity, Population, Export, Cost of Living, Unemployment Rate, Gini

## Briefly introduce the methodology that we are going to use
\qquad According to the question presented, next following things needs to be done: Data check, Conducting the initial EDA(exploratory data analysis) and model check.  
- In the initial EDA part, scatterplots and bubble plot are presented to show the rough corelationship between union density and vocational training to get the general idea of the question.  
-Partial corelation test has been used to check if there still exists significant negative relationship between two variables when control the other 7 confounding variables.    
-Making some adjustment for the model to get more valuable results.  
-Drawing the component residual plot(CR plot) to check if the linear relationship and accordingly make sure that our adjusted models fitted well.  
-To better measure the racial diversity, a new variable based on White/Black/Hispanic is created.  


# Data Check
**Concerns:**
-Sources: Bureau of Labor Statistics
                 wikipedia 
                 U.S. Census Bureau
                 Census Bureau's March Current Population Survey
-Years: from 2016 to 2018
\qquad After looking through the source some issues were found that data is collected from several different source and years, so there comes worries that this data is so effective enough because these variables might have no consistency. It is more helpful to do the data analyze if data is more consistent.

-Variables: Checking 9 variables included in the model, the variable high school stands for the vocational training level. Cause on the state scale, US government needs more fundings to satisfy their supply and demand, so the high school is representitive but it’s better to include more sites such as high technology company, some industries which could contain vocational training people to satisfy their supply and demand as well. It will make more sense to add other potential sites in this variable. Additionally, the variables white/black/hispanic could not show the racial diversity directly,therefore, a new variables based on these three variables is created and will be discussed later in this report.

# Exploratory Data Analysis
- Not a distinguish negative coefficient between vocational training and union density

```{r warning=FALSE, echo=FALSE}
### Initial EDA 
plot1 <-  ggplot(data3)+
  geom_point(aes(x=data3$Highschool,y=data3$Union_density))+
  geom_smooth(aes(x=data3$Highschool,y=data3$Union_density),se=FALSE)+
  labs(x="Union Density",y="Highschool")
plot1
```
\qquad Starting with union density and high school and accordingly reach a scatter plot after cleaning some outliers. While from this plot shows there is not an obvious relationship between these two variables.   

```{r,echo = F}
symbols(data3$Highschool,data3$Union_density,circles= data3$norm_Racial_diversity/pi, inches=0.30, fg="white", bg="lightblue",xlab = "Union Density", ylab = "Highschool")
```
\qquad Moreover, a bubble plot is presented. Same as former plot, seting the union density as independent variable and high school as dependent variable, but this time replacing every spot with a circle, each circle represents for one state and the radius of a circle are standing for racial diversity level. The more racial diversity level, the larger circle is. While, by obverting this plot we could not confirm these two variables has negative relationship either.  

\qquad To better understand this question, a research on the model is conducted and next session will describe this part detailedly.  


# Methodology and Result

### Replicate Original Model

Since the initial EDA does not show a distinctive pattern of the negative relationship between vocational training and unionization, a replication of the original model that is used in the second-year paper has been implemented. With
```{r}
fittest<-lm(Highschool~Union_density+GDP+Urbanity+Population+Export+Unemployment_rate+Cost_of_living+Gini,data = data3)
summary(fittest)
```  

\qquad Since the initial EDA does not show a distinctive pattern of the negative relationship between vocational training and unionization, a replication of the original model that is used in the second-year paper has been implemented. With
```{r,echo = F}
fittest<-lm(Highschool~Union_density+GDP+Urbanity+Population+Export+Unemployment_rate+Cost_of_living+Gini,data = data3)
summary(fittest)
```

### Partial Correlation for Original Model
\qquad Then using the partial correlation test for checking if the negative relationship between Union density and Highschool exists in the primary model. The partial correlation is a measure of the strength and direction of a linear relationship between two continuous variables while controlling for the effect of other continuous variables.  

```{r,echo = F}
n = length(data3$Highschool)
regress.high = lm(Highschool~GDP+Urbanity+Population+Export+Unemployment_rate+Cost_of_living+Gini, data = data3)
regress.Union = lm(Union_density~GDP+Urbanity+Population+Export+Unemployment_rate+Cost_of_living+Gini, data = data3)
residuals.high = residuals(regress.high)
residuals.Union = residuals(regress.Union)
plot(residuals.high,residuals.Union,main="Partial Correlation for Original Model")
abline(lm(residuals.Union~residuals.high), col="red")
```  

\qquad By controlling the seven confounding variables in the primary model, we used two methods on this test. One is the Pearson method, which evaluates the linear relationship between two continuous variables. A relationship is linear when a change in one variable is associated with a proportional change in the other variable. Another is the Spearman method, which evaluates the monotonic relationship between two continuous or ordinal variables. In a monotonic relationship, the variables tend to change together, but not necessarily at a constant rate, and both variables change together does not mean a change in one variable change causes another variable change.  

Pearson Method  

```{r,echo = F}
partial.correlation_p = cor(residuals.high, residuals.Union, method = 'pearson')
test.statistic_p = partial.correlation_p*sqrt((n-2)/(1-partial.correlation_p^2))
p.value_p = 2*pt(abs(test.statistic_p), n-2, lower.tail = F)
partial.correlation_p
p.value_p
```  
Spearman Method  

```{r,echo = F}
partial.correlation_s = cor(residuals.high, residuals.Union, method = 'spearman')
test.statistic_s = partial.correlation_s*sqrt((n-2)/(1-partial.correlation_s^2))
p.value_s = 2*pt(abs(test.statistic_s), n-2, lower.tail = F)
test.statistic_s
p.value_s
```  
\qquad From the output of two methods, those correlations are negative, and P-values are small, which indicates that the negative coefficient of Union density is statistical significance in the primary model. From the plot on residual of Union density and residual of Highschool, the red line has an apparent downward trend. However, by looking at the points on the graph, there is a point in the lower right corner.  



```{r,echo = F}

#check linearity assumption
plot(fittest,which=1)
#check normality assumption
plot(fittest,which=2)
#check constant variance assumption
plot(fittest,which=3)
#check the component residual plot(also called partial residual plot)
crPlots(fittest)
```



- Get the same coefficient
- Model validation

# Transformation
- Use CrPlots to determin which variable to take log transformation
- Model validation



```{r,echo = F}

fit1<-lm(log(Highschool) ~ Union_density+log(GDP)+Urbanity+log(Population)+log(Export)+Unemployment_rate+log(Cost_of_living)+Gini,data = data3)
summary(fit1)
crPlots(fit1)
plot(fit1,which=1)
plot(fit1,which=2)
plot(fit1,which=3)
```  

\qquad Because the original model does not in line with the assumption, data transformation is used to adjust the model. To use the linear regression model, the most significant assumption is that variables should be in the linear relationship with the outcome. Component residual plot is used to measure if every variable satisfies linearity assumption. The red line shows the pattern of the datapoints included in every variable, and the dashed line indicates the best of the fit. If the red line is close to the dashed line and does not show curve pattern, meaning that the linearity assumption is satisfied for that variable. Based on the initial component residual plot of the original model, leverage points exist in datapoints from different variables including “GDP”, “population”, “Export”, “unemployment rate” and “cost of living”. There are issues with leverage points since it would directly change the pattern of the residual plots. For example, looking at the residual plots of “Export”, if the leverage points move upwards or downwards, the red line and the dashed line would move accordingly. The movement is not caused by the movement of the most datapoints, rather caused by the leverage points, which is not the ideal case where the red line should indicate the pattern of the majority of datapoints. To avoid the leverage points, log transformation is used to build the first model, “fit1”. 
In model “fit1”, log transformation has been implemented for the outcome, “highschool” and variables that have leverage points, “GDP”, “population”, “Export”, “unemployment rate” and “cost of living”. After summarizing the model “fit1”, the coefficient between the outcome “highschool” and the variable “union density” is -0.04, not significant. Because the dataset is a sample from the population, coefficient for each sample would be different because of the sampling bias arising from picking up different samples within the population. The insignificant coefficient indicates that the negative relationship between vocational training and union density is not conclusive.


\qquad After modeling, residual plots are generated to check the validation of the model. Looking at the residual plot, nearly all the residuals are equally distributed within the horizontal line valued at zero, indicating that the residuals are random, and the linearity assumption is satisfied. Looking at the Q-Q plot, majority of the residual points are aligned with the straight line, indicating the normality assumption is nearly satisfied. Looking at the scale-location plot, most residues are spread equally along the ranges of predictors, indicating the equal variance assumption is satisfied. Based on the residual plot, the model “fit1” is more valid compared to the original model. Check the component and residual plot again, the issue regarding the leverage points have been improved a lot. But two issues still have to be considered. First, there is a leverage point in GDP after taking log transformation. Second, the variable of union density slightly shows a curved pattern, indicating that linear assumption is not well satisfied, and a more robust model might be considered.

```{r}
data4 <-data3 %>% 
  filter(GDP<100000)
fit2<-lm(log(Highschool) ~ Union_density+log(GDP)+Urbanity+log(Population)+log(Export)+Unemployment_rate+log(Cost_of_living)+Gini,data = data4)
summary(fit2)
crPlots(fit2)
plot(fit2,which=1)
plot(fit2,which=2)
plot(fit2,which=3)
```

\qquad To deal with the leverage point in log GDP, one of the solutions is to leave it out. The leverage point is Washington DC with highest log GDP. Another model (fit2) with exactly same variables and outcome is created with a new dataset excluding datapoint of Washington DC. After summarizing the model “fit2”, the coefficient of union density changes 20%, indicating the difference is not ignorable.Thus, the leverage point of Washington DC cannot be simply leaved out.

### Partial Correlation

# Adding Racial Diversity

```{r}
fit3<-lm(log(Highschool) ~ Union_density+log(GDP)+Urbanity+log(Population)+log(Export)+Unemployment_rate+log(Cost_of_living)+Gini + norm_Racial_diversity,data = data3)
summary(fit3)
plot(fit3,which=1)
plot(fit3,which=2)
plot(fit3,which=3)
```

### Mediation Analysis

# Conclusion


